A few stickleback suffice to transport adaptive alleles to new lakes

### 2018-12-11 Stickleback_Paper.tex
Jared's edits to the introduction 2

The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations. 
While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures. 
A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.
%Which evolutionary processes promote the maintenance of re-use of such genetic polymorphisms? 
%How do these processes interact with diverse geographic contexts, as well as variation in populations genetic processes such as gene flow, recombination, selection to fully take advantage of SGV to locally adapt. 
However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
How do complex geographic systems with a variety of selective pressures interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations?

Recently, a large flood of data resulting from systems such as this come from population genomic studies in nature that have been powered by advances in sequencing technologies. 
%One such organism is the Threespine Stickleback.
One such organism is the Threespine Stickleback which has exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
%This small fish has long been used for studies of behavior, ecology and evolution in the wild, and has more recently become a model for understanding the genetic basis of adaptive evolution. 
%The ancestral marine form has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
The rigidity of the species' ability to frequently prosper in newly created freshwater environments has made this fish a good model for understanding the genetic basis of adaptive evolution. 
The ancestral marine form of Stickleback has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
%which have each evolved a similar suite - or 'freshwater syndrome' - of phenotypes despite local differences. 
%Phnotypically, these freshwater patches which have 
Geographic isolation prevents direct migration between a large majority of these patches of freshwater populations,
yet we observe very similar phenotypes and many shared adaptive alleles often found to be identical by decent (IBD). 
%Impressively, this adaptive phenotypic evolution to freshwater environments can occur in mere millenia, in many cases since the end of the last ice age about 13,000 years ago. 
%Studies of the fossil record also clearly show that invasion of oceanic stickleback and local adaptation in freshwater has occurred repeatedly over millions of years, 
%often involving very similar phenotypic shifts as seen in contemporary populations.
Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, 
and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago (Cite Kristin).
%A finding from recent stickleback population genomic studies is that the same genomic regions, and in some cases the same haplotypes, are implicated in patterns of repeated evolution. (Cite?)
%These studies show that divergent stickleback populations are clearly still connected via bi-directional gene flow. 
%This reuse of standing genetic variation motivated empirical studies to ask whether this phenotypic and genomic evolution could occur in decades. 
%The answer is a clear yes. 
For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island. 
Quickly inhabited by the surrounding marine population of stickleback, 
\citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years. 
In these freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows
the same expansion of the dorsal region and reduction of the ventral region as observed in the large majority of freshwater demes.

But how can evolution occur at such a rapid pace? 
Waiting for new mutations to arise in each lake would take much longer, and genotypes being identical by state is even more improbable. 
This would seem to suggest that freshwater alleles are maintained in the marine individuals by repeated adaptation to freshwater environments and downstream gene flow from those locally adapted sub-populations.
%This suggests that adaptation to freshwater environments is accelerated by selection on SGV found in marine populations. 
The first clear example of the global reuse of SGV was the gene \textit{eda} which has been shown to be an important regulator for the number of lateral plates. 
While the low lateral plate version of this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently. 
More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}. 
%Genomic regions that exhibit local adaptation, as evidenced by increased relative population genomic divergence, also show much older haplotypes that are enriched in the alternative habitats.

>>>>

The canonical model for the genetics of adaptation, first formulated by Fisher in the early part of the 20th century, involves the sequential fixation of new mutations. 
While it has proven valid in numerous studies in the field as well as in the lab, this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures. 
A growing number of studies have identified instances of convergent or shared adaptive evolution in small demes that utilize standing genetic variation (SGV) found in the static population from which they split.
However, it is still widely unknown which variations of evolutionary processes such as gene flow, recombination, selection and mutation, promote the maintenance and re-use of such genetic polymorphisms.
How do complex geographic systems with a variety of selective pressures interact with these evolutionary forces to help or hinder the use of SGV in newly colonized subpopulations?

Recently, a large flood of data resulting from systems such as this come from population genomic studies in nature that have been powered by advances in sequencing technologies. 
One such organism is the Threespine Stickleback which has exhibited a long standing evolutionary history of migration-selection balance between marine and freshwater populations across the globe.
The rigidity of the species' ability to frequently prosper in newly created freshwater environments has made this fish a good model for understanding the genetic basis of adaptive evolution. 
The ancestral marine form of Stickleback has given rise to millions of independently derived populations in recently de-glaciated regions of the Northern Hemisphere.
Geographic isolation prevents direct migration between a large majority of these patches of freshwater populations,
yet we observe very similar phenotypes and many shared adaptive alleles often found to be identical by decent (IBD). 
Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in tens of generations, 
and the adaptive alleles are found in freshwater populations that have been geographically isolated since the end of the last ice age $\approx 13,000$ years ago (Cite Kristin).
For example, in 1964 the Great Alaskan Earthquake caused an uplift of Middleton island and in turn, introduced a group of freshwater ponds around the perimeter of the island. 
Quickly inhabited by the surrounding marine population of stickleback, 
\citet{lescak2015evolution} observed significant phenotypic changes in less than 50 years that appear to be parallel to freshwater stickleback that have been separated for over thousands of years. 
In these freshwater stickleback, the number of lateral plates are reduced and the opercle shapes shows
the same expansion of the dorsal region and reduction of the ventral region as observed in the large majority of freshwater demes.

But how can evolution occur at such a rapid pace? 
Waiting for new mutations to arise in each lake would take much longer, and genotypes being identical by state is even more improbable. 
This would seem to suggest that freshwater alleles are maintained in the marine individuals allowing the accelerated selection on SGV found in marine individuals which colonize the lake. 
%by repeated adaptation to freshwater environments and downstream gene flow from those locally adapted sub-populations.
%This suggests that adaptation to freshwater environments is accelerated by selection on SGV found in marine populations. 
The first clear example of the global reuse of SGV was the gene \textit{eda} which has been shown to be an important regulator for the number of lateral plates. 
While the low lateral plate version of this gene arose millions of years ago, it is found in freshwater ponds which have formed much more recently. 
More recently, population genomic studies employing genome-wide haplotype analyses has provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}. 
%Genomic regions that exhibit local adaptation, as evidenced by increased relative population genomic divergence, also show much older haplotypes that are enriched in the alternative habitats.
